
# Vector functions

```{r message = FALSE, warning = FALSE}
library(tidyverse)
```


**Vector functions** take a vector as input and produce a vector of the same length as output. This is very helpful when working with vectors. For example, instead of taking the log of each element of the vector `x`, you can just call `log10()` once:

```{r}
x <- c(5, 2, 1)

log10(x)
```

The simple mathematical operators are also vector functions:

```{r}
y <- c(1, 2, 4)

x + y

x * y
```

In contrast, functions that can only take a length one input and produce a length one output are called **scalar functions**. 

As you'll see in the next section, the distinction between scalar and vector functions is important when working with tibbles.

## Temperature recommendations

A common way to create a scalar function is by using an if-else statement. For example, you might write the following function that tells you what to do based on the temperature outside:

```{r}
recommendation_1 <- function(x) {
  if (x >= 90) {
    "locate air conditioning"
  } else if (x >= 60) {
    "go outside"
  } else if (x >= 30) {
    "wear a jacket"
  } else if (x >= 0) {
    "wear multiple jackets"
  } else {
    "move"
  }
}
```

This works well when applied to single values:

```{r}
recommendation_1(92)

recommendation_1(34)

recommendation_1(-15)
```

but fails when applied to a vector with more than one element:

```{r}
temps <- c(1, 55, 101)

recommendation_1(temps)
```

`if` only works with one element at a time and can't handle an entire vector. When you give `recommendation_1()` a vector, it only processes the first element of that vector, which is why `recommendation_1()` only tells us what to do if it's 1 degree outside.

## Vector functions and `mutate()`

`mutate()` creates a value for each row in a tibble. If you want, you can manually give `mutate()` a vector with a value for each row:

```{r}
set.seed(523)

df <- tibble(
  temperature = sample(x = -15:110, size = 10, replace = TRUE)
)

df %>% 
  mutate(new_column = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10))
```

You can also give `mutate()` a single value:

```{r}
df %>% 
  mutate(one_value = 1)
```

and it will repeat that value for each row in the tibble. However, if you try to give `mutate()` a vector with a length other than 1 or `nrow(df)`, you'll get an error:

```{r, error=TRUE}
df %>% 
  mutate(two_values = c(1, 2))
```

As you know well by now, you'll often create new columns by applying functions to existing ones:

```{r}
fahrenheit_to_celcius <- function(degrees_fahrenheit) {
  (degrees_fahrenheit - 32) * (5 / 9)
}

df %>% 
  mutate(temperature_celcius = fahrenheit_to_celcius(temperature))
```

When you pass `temperature` to `fahrenheit_to_celcius()`, you pass the entire `temperature` column, which, as you learned earlier, is a vector. Because mathematical operations are vectorized, `fahrenheit_to_celcius()` returns a vector of the same length and `mutate()` successfully creates a new column.

You can probably predict now what will happen if we try to use our scalar function, `recommendation_1()`, in the same way:

```{r}
df %>% 
  mutate(recommendation = recommendation_1(temperature))
```

`mutate()` passes the entire `temperature` vector to `recommendation_1()`, which can't handle a vector and so only processes the first element of `temperature`. However, because of how `mutate()` behaves when given a single value, the recommendation for the first temperature is copied for every single row, which isn't very helpful. 

## Vectorizing if-else statements

There are several ways to vectorize `recommendation_1()` so that it gives an accurate recommendation for each temperature in `df`.

First, there's a vectorized if-else function called `if_else()`:

```{r}
x <- c(1, 3, 4)

if_else(x == 4, true = "four", false = "not four")
```

However, in order to rewrite `recommendation_1()` using `if_else()`, we'd need to nest `if_else()` repeatedly and the function would become difficult to read. Another vector function, `case_when()`, is a better option:

```{r}
recommendation_2 <- function(x) {
  case_when(
    x >= 90 ~ "locate air conditioning",
    x >= 60 ~ "go outside",
    x >= 30 ~ "wear a jacket",
    x >= 0  ~ "wear multiple jackets",
    TRUE    ~ "move"
  )
}

recommendation_2(temps)

df %>% 
  mutate(recommendation = recommendation_2(temperature))
```

For other helpful vector functions, take a look at the "Vector Functions" section of the [dplyr cheat sheet](https://github.com/rstudio/cheatsheets/blob/master/data-transformation.pdf).
