
# Tidy evaluation basics

```{r setup, include=FALSE}
library(tidyverse)
```

Likely without realizing it, you've been using two types of variables in R. First, there are what we'll call _environment_ variables. When you create an environment variable, you bind a value to a name in the current environment. For example, the following code creates an environment variable named `y` that refers to `1`.

```{r}
y <- 1
```

Now, when we call `y`, we get whatever value `y` refers to. 

```{r}
y
```

The second type of variable we'll call _data_ variables. Data variables refer to columns in data frames, and only make sense in the context of those data frames. For example, `manufacturer` is a data variable that exists in the context of `mpg`. dplyr, ggplot2, and other tidyverse packages understand this, which is why functions like `select()` behave as you want.

```{r}
mpg %>% 
  select(manufacturer)
```

Instead of figuring out what `manufacturer` refers to in the environment, `select()` looks for a column named "manufacturer" in `mpg`. 

Let's take a look at our failed function from the intro again.

```{r, error=TRUE}
grouped_mean <- function(var_group, var_summary) {
  mpg %>% 
    group_by(var_group) %>% 
    summarize(mean = mean(var_summary))
}

grouped_mean(var_group = manufacturer, var_summary = cty)
```

`group_by()` thinks that `var_group` is a data variable, and so it looks inside `mpg` for a column called "var_group", doesn't find it, and so throws an error. 

We actually want `var_group` to behave as a hybrid between an environment variable and a data variable. Like an environment variable, we want it to refer to another value (`manufacturer`). Then, we want `group_by()` to treat that value (`manufacturer`) as a data variable and look inside `mpg` for the matching column.

## Embracing with `{{ }}`

To fix `grouped_mean()`, all we need to do is _embrace_ `var_group` and `var_summary` with `{{ }}` (pronounced "curly curly").

```{r}
grouped_mean <- function(var_group, var_summary) {
  mpg %>% 
    group_by({{ var_group }}) %>% 
    summarize(mean = mean({{ var_summary }}))
}

grouped_mean(var_group = manufacturer, var_summary = cty)
```

`{{ }}` might remind you of the use of `{ }` in `glue()`, and they fill similar roles. 

Let's look at another example. This doesn't work:

```{r, error=TRUE, fig.show='hide'}
plot_mpg <- function(var) {
  mpg %>% 
    ggplot(aes(var)) +
    geom_bar()
}

plot_mpg(drv)
```

We want `var` to behave as a hybrid environment-data variable, so we need `{{ }}`.

```{r}
plot_mpg <- function(var) {
  mpg %>% 
    ggplot(aes({{ var }})) +
    geom_bar()
}

plot_mpg(drv)
```

## Forwarding with `...`

You might have noticed that some functions, like the purrr functions, take `...` as a final argument, allowing you to specify any number of additional arguments. You can use `...` in your own functions. There are two common use-cases.

### Passing full expressions

Functions like `filter()` take expressions, like `year == 1999` or `manufacturer == "audi"`. If you want to build a function that takes full expressions, you can use `...`.

```{r}
mpg_filter <- function(...) {
  mpg %>% 
    filter(...)
}

mpg_filter(manufacturer == "audi", year == 1999)
```

`...` can take any number of arguments, so we can filter by an unlimited number of conditions.

```{r}
mpg_filter(
  manufacturer == "audi", 
  year == 1999, 
  drv == "f", 
  fl == "p"
)
```

`mpg_filter()` _forwards_ `...` to `filter()`, which allows `filter()` to act on the contents of `...` just as it would outside of the function.

Here's another example that uses `select()`.

```{r}
mpg_select <- function(...) {
  mpg %>% 
    select(...)
}

mpg_select(car = model, drivetrain = drv)
```

### Additional arguments

Sometimes, you'll want your function to take named arguments, but you'll also want to allow for any number of additional arguments. You can use `{{ }}`, and `...`.

```{r}
grouped_mean_2 <- function(df, var_summary, ...) {
  df %>% 
    group_by(...) %>% 
    summarize(mean = mean({{ var_summary }})) %>% 
    ungroup()
}

grouped_mean_2(df = mpg, var_summary = cty, year, drv)
```

With the `...`, we can pass any number of grouping variables into `group_by()`.

```{r}
grouped_mean_2(df = mpg, var_summary = cty, year, drv, class)
```

## Assigning names

When you want to pass the name of a column into your function, you need to:

* Embrace the name with `{{ }}`.
* Use `:=` instead of `=` to assign the name.

```{r}
summary_mean <- function(df, var_summary, name_summary){ 
  df %>% 
    summarize({{ name_summary }} := mean({{ var_summary }}))
}

summary_mean(df = mpg, var_summary = cty, name_summary = cty_mean)
```

You have to use `:=` instead of just plain `=` because you can't use `{{ }}` on both sides of a `=`.

(`:=` is called the walrus operator because it looks like a sideways walrus.)

## Splicing with `!!!`

Say you want to recode a variable:

```{r}
mpg %>% 
  mutate(drv = recode(drv, "f" = "front", "r" = "rear", "4" = "four"))
```

It's often a good idea to store your recode mapping as a vector in your parameters section. To get this to work, you'll need another tidyeval operator, `!!!`. 

```{r}
recode_drv <- c(f = "front", r = "rear", "4" = "four")

mpg %>% 
  mutate(drv = recode(drv, !!! recode_drv))
```

`!!!` has two tasks:

* Unpack `recode_drv` so that each element is passed as a separate argument to `recode()` (i.e., `"f" = "front", "r" = "rear", "4" = "four"` instead of c(`"f" = "front", "r" = "rear", "4" = "four")`).
* Make sure that `recode()` treats the individual elements as data variables. 

