
# Summary

```{r setup, include=FALSE}
library(tidyverse)
```

Tidy evaluation is a complex subject, but you don't really need to understand how it works to start using it i your functions. The following is a brief, practical summary of the most common use-cases. 

## Passing full expressions with `...`

You can use `...` to pass full expressions into a function. For example, the following function applies `filter()` to `mpg.`

```{r}
mpg_filter <- function(...) {
  mpg %>% 
    filter(...)
}

mpg_filter(manufacturer == "audi", year == 1999)
```

`...` can take any number of arguments, so we can filter by an unlimited number of conditions.

```{r}
mpg_filter(
  manufacturer == "audi", 
  year == 1999, 
  drv == "f", 
  fl == "p"
)
```

Passing `...` works anytime all you want to do is pass a full expression into a dplyr function. Here's another example that uses `select()`.

```{r}
mpg_select <- function(...) {
  mpg %>% 
    select(...)
}

mpg_select(car = model, drivetrain = drv)
```

## Named arguments

Sometimes, `...` won't work because you'll want to supply your function with named arguments.

In the introduction, we wrote the following function that doesn't work:

```{r error=TRUE}
grouped_mean <- function(df, group_var, summary_var) {
  df %>% 
    group_by(group_var) %>% 
    summarize(mean = mean(summary_var))
}

grouped_mean(df = mpg, group_var = manufacturer, summary_var = cty)
```

Use `enquo()` and `!!` to create a function that works. 

```{r}
grouped_mean <- function(df, group_var, summary_var) {
  group_var <- enquo(group_var)
  summary_var <- enquo(summary_var)
  
  df %>% 
    group_by(!!group_var) %>% 
    summarize(mean = mean(!!summary_var))
}

grouped_mean(df = mpg, group_var = manufacturer, summary_var = cty)
```

Here's the steps:

* Apply `enquo()` to the arguments that refer to column names.
* When you want to reference those arguments, put `!!` before their names.

## Named arguments and any number of additional arguments

Sometimes, you'll want your function to take named arguments, but you'll also want to allow for any number of additional arguments. You can use `enquo()`, `!!`, and `...`.

```{r}
grouped_mean_2 <- function(df, summary_var, ...) {
  summary_var <- enquo(summary_var)
  
  df %>% 
    group_by(...) %>% 
    summarize(mean = mean(!!summary_var))
}

grouped_mean_2(df = mpg, summary_var = cty, year, drv)
```

With the `...`, we can pass any number of grouping variables into `group_by()`.

```{r}
grouped_mean_2(df = mpg, summary_var = cty, year, drv, class)
```

## Assigning names

Many of the dpylr verbs allow you to name or rename columns. When you want to pass the name of a column into your function, you need to:

* Apply `enquo()` to the argument giving the name
* Put `!!` before the name when you reference it
* Use `:=` instead of `=` to assign the name

```{r}
summary_mean <- function(df, summary_var, summary_name) {
  summary_var <- enquo(summary_var)
  summary_name <- enquo(summary_name)
  
  df %>% 
    summarize(!!summary_name := mean(!!summary_var))
}

summary_mean(df = mpg, summary_var = cty, summary_name = cty_mean)
```

## Recoding

`recode()` is also a dplyr verb. It is often useful to put your recode vector in the parameters section, instead of directly inside `recode()`. Use `!!!` to get this to work.

```{r}
recode_drv <- c("f" = "front", "r" = "rear", "4" = "four")

mpg %>% 
  mutate(drv = recode(drv, !!!recode_drv))
```

