
# Tidy evaluation

```{r message = FALSE, warning = FALSE}
library(tidyverse)
```

Now, we'll go into some more detail about why you need tidy evaluation and explain what's happening when you call `enquo()` or use `!!`.

## When do you need tidy evaluation?

You don't always need tidy evaluation when programming with dplyr or other tidyverse functions that work with tibbles. For example, the following function works just fine.

```{r}
filter_fun_1 <- function(df, value) {
  df %>% 
    filter(model == value) %>% 
    nrow()
}

filter_fun_1(df = mpg, value = "corvette")
```

We can make this function more general by adding an argument to specify which column to supply to `filter()`. However, the following function doesn't work.

```{r error=TRUE}
filter_fun_2 <- function(df, var, value) {
  mpg %>% 
    filter(var == value) %>% 
    nrow()
}

filter_fun_2(var = model, value = "corvette")
```

You need tidy evaluation if you want to build a function that passes the names of tibble columns into dplyr verbs.

Here's another example of a function that doesn't work:

```{r error=TRUE}
grouped_mean <- function(df, group_var, summary_var) {
  mpg %>% 
    group_by(group_var) %>% 
    summarize(mean = mean(summary_var))
}

grouped_mean(mpg, manufacturer, cty)
```

In the following sections, we'll explore why `mpg_grouped_mean()` doesn't work and show you how to create a function that does.

## Evaluated vs. quoted arguments

Before we can explain what's going wrong in functions like `grouped_mean()`, we need to lay some groundwork. In R, you can divide function arguments into two classes: __evaluated__ and __quoted__. 

Evaluated arguments are what you might think of as "normal." Here's an example of a function that evaluates its function arguments.

```{r}
log(2)
```

We can refer to 2 in any way we want and get the same answer.

```{r}
x <- 2

log(x)
```

Code in an evaluated argument executes the same regardless of whether or not its in a function argument. So `x` refers to 2 whether it's outside `log()`

```{r}
x
```

or inside.

```{r}
log(x) == log(2)
```

Because `log()` evaluates its arguments, it figures out what `x` refers to before operating on `x`. In our example, `log()` figures out that `x` refers to 2, and then takes the log of 2. This might seem obvious to you, because what would it even mean to take the log of `x` if `x` didn't refer to a number? Sometimes, however, functions don't want to evaluate their arguments. 

Let's find out if dplyr functions use evaluated arguments. Here's a dplyr function.

```{r}
mpg %>% 
  select(cty)
```

In this case, let's just consider the `cty` argument, even though `mpg` is technically an argument as well. 

If `select()` evaluates its argument, then `cty` should refer to the same thing inside `select()` as it does outside `select()`. Inside `select()`, we know that `cty` refers to a column of integers. Does it refer to the same outside `select()`?

```{r, error=TRUE}
cty
```

No. R doesn't know what `cty` refers to, because `cty`, unlike `x`, doesn't exist in the global environment. It only exists as a column of `mpg`. 

To make this point even more explicit, let's assign a random number to `cty`.

```{r}
cty <- 3

cty
```

Now, outside `select()`, `cty` refers to `3`, but still refers to the tibble column inside `select()`.

```{r}
mpg %>% 
  select(cty)
```

Arguments like `cty` are __quoted__. Instead of immediately evaluating `cty` before operating, `select()` and the other dplyr verbs hold on to what was literally supplied as an argument. Here, that's just "cty", but in other situations might be more complicated. Quoting their arguments allow `select()` and the other dplyr verbs to use their input as they want, without worrying about how they would evaluate in the global environment. In our example, quoting its argument allows `select()` to look for `cty` inside the `mpg` tibble, without worrying about what `cty` may or may not refer to in the global environment (like the number 3).

Note that dplyr verbs only quote the arguments that have to do with column names. They do not quote the argument that refers to the data you pipe in, or non-column-name arguments like `count()`'s `sort` argument or `top_n()`'s `n` argument.

dplyr's quoting behavior makes it really easy to use. You can supply bare names of columns to dplyr functions and they just work. However, the quoting behavior causes some wrinkles when you want to program with dplyr. 

You might have a hypothesis now about why our `grouped_mean()` function didn't work. Here it is again.

```{r, error=TRUE}
grouped_mean <- function(df, group_var, summary_var) {
  df %>% 
    group_by(group_var) %>% 
    summarize(mean = mean(summary_var))
}

grouped_mean(df = mpg, group_var = manufacturer, summary_var = cty)
```

The error is that the column `group_var` is unknown. Because `group_by()` quotes its argument, it didn't evaluate `group_var`, find out that it refers to `manufacturer`, and then look for a column named `manufacturer`. Instead, it took its input literally and looked for a column named `group_var`. `mpg` doesn't have a column called `group_var`, so we got an error.

We `group_by()` to understand that `group_var` refers to `manufacturer`, so we're going to have to make some changes to our function. Before diving into the specifics of these changes, here's a summary of the points covered so far:

* Some functions evaluate their arguments and some function quote their arguments.
* You can tell if an argument is quoted if it produces different results inside and outside of the function.
* dplyr functions quote their arguments (the ones that have to do with tibble columns).

### Strings and `glue()`

If we want `group_by()` to understand that `group_var` refers to `manufacturer`, we're going to have to _unquote_ `group_var`. 

Before we talk about how to do this with dplyr, let's take a moment to examine a situation in which you've actually already been quoting and unquoting input. 

You already know that, when you create a string, R doesn't evaluate their content. When you type something like:

```{r}
"y <- 1"
```

R creates a string, not an object named `y` with a value of 1. 

Sometimes, though, you'll want to write a function that inserts a variable into a string. For example, say we want to write a function that tells you what species you are. 

You can probably predict that the following function won't work.

```{r}
species <- function(my_species) {
  "I am a my_species"
}

species("human")
```

We need to tell our function that we actually do want to evaluate `my_species`. As you've already learned, you can do this with `str_glue` and `{}`.

```{r}
species <- function(my_species) {
  str_glue("I am a {my_species}")
}

species("human")
```

The `{}`s tell `str_glue()` to evaluate `my_species` before constructing the string. 

This situation is similar to our `mpg_grouped_mean()` conundrum. Unfortunately, we can't just use `{}` to get dplyr verbs to evaluate the arguments we want. We'll need to figure out an equivalent to to `{}` inside dplyr function calls.

### Quosures 

One reason we can't just generalize from our string example and use `{}` is that when dplyr functions quote their arguments, they don't quote and create strings. When you quote with quotation marks, as you know, you create a string.

```{r}
"y <- 1"
```

But when dplyr functions quote their arguments, they create something called a _quosure_.

Let's create a quosure to try to figure out what exactly quosures are. You can create your own quosure with the function `quo()`. 

```{r}
quo(y <- 1)
```

Notice that the quosure has two parts: the `expr` (which stands for "expression") and the `env` (which stands for "environment"). 

You can think of _expressions_ like recipes. A recipe for chocolate chip cookies specifies how to make the cookies, but does not itself create any cookies. Similarly, the expression `y <- ` species how to create a variable, but doesn't actually create that variable. Just like you to carry out the recipe to create cookies, R needs to evaluate the expression to produce the results.

Recipes, unfortunately, aren't sufficient for cookies. You also need a stock of ingredients, like flour and chocolate chips. Similarly, in order to evaluate an expression, R needs an environment that supplies the variables. Different types of flour and chocolate chips can create different cookies, and different environments can cause the same expression to be evaluated differently.

If we place `quo(y <- 1)` inside a function, the environment will change.

```{r}
quo_fun <- function() {
  quo(y <- 1)
}

quo_fun()
```

Quosures are objects and so can be passed around, carrying their environment with them.

```{r}
more_quo_fun <- function(my_quosure) {
  my_quosure
}

more_quo_fun(quo(y <- 1))
```

Now you know that:

* dplyr quotes its arguments and creates quosures, which consist of an expression and an environment. An expression is kind of like a recipe, and the environment is what supplies the ingredients you use to carry out that recipe.
* You can create a quosure with `quo()`.

## Wrapping quoting functions

### `enquo()` and `!!`

Now that we've gone over some theory, we can return to the task of fixing `mpg_grouped_mean()`.

You just learned that dplyr verbs create quosures. In order to make our function work, we'll need to make our own quosure that captures our desired meaning of `group_var` and `summary_var`. Here's our earlier, unsucessful function.

```{r, error=TRUE}
grouped_mean <- function(df, group_var, summarize_var) {
  df %>% 
    group_by(group_var) %>% 
    summarize(mean = mean(summarize_var))
}

grouped_mean(df = mpg, group_var = manufacturer, summarize_var = cty)
```

We want to track the environment of `group_var` so that `group_by()` knows that it refers to `manufacturer`. We can do this with `quo()`.

```{r, error=TRUE}
grouped_mean <- function(df, group_var, summarize_var) {
  print(group_var)
  
  df %>% 
    group_by(group_var) %>% 
    summarize(mean = mean(summarize_var))
}

grouped_mean(
  df = mpg, 
  group_var = quo(manufacturer), 
  summarize_var = quo(cty)
)
```

Our `print()` statement lets us know what `group_var` looks like inside the function. `group_var` is a quosure and evaluates to `manufacturer,` which seems like a step in the right direction. However, our function still isn't giving us what we want. `group_by()` is still quoting its argument and looking for a column called `group_var` because we haven't given `group_by()` any indication that it shouldn't be quoting. `group_by()` doesn't know we've already done the quoting (with `quo()`), and so it's just carrying on normally. 

We can tell `group_by()` not to quote by using `!!` (pronounced "bang bang"). `!!` says something like "evaluate me!" or "unquote!"

```{r}
grouped_mean <- function(df, group_var, summarize_var) {
  print(group_var)
  
  df %>% 
    group_by(!! group_var) %>% 
    summarize(mean = mean(!! summarize_var))
}

grouped_mean(
  df = mpg, 
  group_var = quo(manufacturer), 
  summarize_var = quo(cty)
)
```

Success!!

`quo()` and `!!` work well, but it's kind of a hassle to have to `quo()` our input each time. It would be even better if we could write the funcion call like this:

```{r, eval=FALSE}
grouped_mean(df = mpg, group_var = manufacturer, summarize_var = cty)
```

To do so, we'll take care of the quoting _inside_ our function. We can't use `quo()` to quote inside our function.

```{r, error=TRUE}
grouped_mean <- function(df, group_var, summary_var) {
  group_var <- quo(group_var)
  summary_var <- quo(summary_var)
  print(group_var)
  
  df %>% 
    group_by(!! group_var) %>% 
    summarize(mean = mean(!! summary_var))
}

grouped_mean(df = mpg, group_var = manufacturer, summary_var = cty)
```

The environment of our quosure is wrong. We want R to evaluate `group_var` using the global environment, not the environment of our function. We'll need `quo()`'s cousin, `enquo()`, in order to capture the correct environment of `group_var`. 

```{r}
enquo_fun <- function(group_var) {
  print(enquo(group_var))
}

enquo_fun(group_var = manufacturer)
```

Now, we can rewrite `mpg_grouped_mean()`.

```{r}
grouped_mean <- function(df, group_var, summary_var) {
  group_var <- enquo(group_var)
  summary_var <- enquo(summary_var)

  df %>% 
    group_by(!! group_var) %>% 
    summarize(mean = mean(!! summary_var))
}

grouped_mean(df = mpg, group_var = manufacturer, summary_var = cty)
```

You can use this same technique for any of the dplyr verbs.

```{r}
filter_var <- function(var, value) {
  var <- enquo(var)
  
  mpg %>% 
    filter(!! var == value)
}

filter_var(class, "minivan")
```

The `enquo()` and `!!` strategy is incredibly useful, and you don't need to fully understand the theory behind it in order to write successful functions. If some of the earlier explanation is still confusing, don't worry about it too much. Tidy evaluation is a complicated subject, and it takes a while to really grasp what's going on behind the `enquo()` and `!!`'s.

In summary, to build a function that takes an argument to a dplyr verb, use the following template:

```{r eval=FALSE}
my_tidyeval_function <- function(column_name) {
  column_name <- enquo(column_name)
  
  df %>% 
    dplyr_verb(!! column_name)
}
```

### Passing `...`

Say you want to extend `mpg_grouped_mean()` so that you can group by any number of variable. You might have noticed that some functions, like scoped verbs and the purrr functions, take ... as a final argument, allowing you to specify additional arguments to their functions. We can use that same functionality here.

```{r}
grouped_mean_2 <- function(df, summary_var, ...) {
  summary_var <- enquo(summary_var)
  
  df %>% 
    group_by(...) %>% 
    summarize(mean = mean(!! summary_var))
}

grouped_mean_2(df = mpg, summary_var = cty, manufacturer, model)
```

Notice that with ..., we didn’t have to use enquo() or !!. ... takes care of all the quoting and unquoting for you.

You can also use ... to pass in full expressions to dplyr verbs.

```{r}
filter_fun <- function(df, summary_var, ...) {
  summarize_var <- enquo(summary_var)
  
  df %>% 
    filter(...) 
}

filter_fun(mpg, manufacturer == "audi", model == "a4")
```

## Assigning names

Let's return to our `mpg_grouped_mean()` function. We finally got it working in the last section. Here is it again:

```{r}
grouped_mean <- function(df, group_var, summarize_var) {
  group_var <- enquo(group_var)
  summarize_var <- enquo(summarize_var)
  
  df %>% 
    group_by(!! group_var) %>% 
    summarize(mean = mean(!! summarize_var))
}
```

It would be nice if we could name the `mean` column something more informative than just `mean`.

Maybe we can just apply our `enquo()` and `!!` strategy and call it a day?

```{r eval=FALSE}
grouped_mean <- function(df, group_var, summary_var, summary_name) {
  group_var <- enquo(group_var)
  summary_var <- enquo(summary_var)
  summary_name <- enquo(summary_name)
  
  df %>% 
    group_by(!! group_var) %>% 
    summarize(!! summary_name = mean(!! summary_var))
}

grouped_mean(
  df = mpg, 
  group_var = manufacturer, 
  summary_var = hwy, 
  summary_name = mean_hwy
)
```

Sadly, this doesn't work. It turns out that you can't use `!!` on both sides of an `=`. We have to use a special `=` that looks like `:=`.

```{r}
grouped_mean <- function(df, group_var, summary_var, summary_name) {
  group_var <- enquo(group_var)
  summary_var <- enquo(summary_var)
  summary_name <- enquo(summary_name)
  
  df %>% 
    group_by(!! group_var) %>% 
    summarize(!! summary_name := mean(!! summary_var))
}

grouped_mean(
  df = mpg, 
  group_var = manufacturer, 
  summary_var = hwy, 
  summary_name = mean_hwy
)
```

Success!!

## Passing vectors with `!!!`

Here's one more useful tidy evaluation use-case. 

Say you want to use `recode()` to recode a variable. 

```{r}
mpg %>% 
  mutate(drv = recode(drv, "f" = "front", "r" = "rear", "4" = "four")) %>% 
  select(drv)
```

It's often a good idea to store you recode mapping in a parameter. You might want to change the mapping later on, or use it in other locations. 

We can store the mapping in a named character vector.

```{r}
drv_recode <- c("f" = "front", "r" = "rear", "4" = "four")
```

However, now `recode()` doesn't work.

```{r, error=TRUE}
mpg %>% 
  mutate(drv = recode(drv, drv_recode)) %>% 
  select(drv)
```

`recode()`, like `group_by()`, `summarize()`, and the other dplyr functions, quotes its input. We therefore need to tell `recode()` evaluate `recode_key` immediately. Let's try `!!`.

```{r, error=TRUE}
mpg %>% 
  mutate(drv = recode(drv, !!drv_recode)) %>% 
  select(drv)
```

`!!` doesn't work because `recode_key` is a vector. Not only do we need to immediately evaluate `recode_key`, we also need to unpack its contents. To do so, we'll use `!!!`.

```{r}
mpg %>% 
  mutate(drv = recode(drv, !!!drv_recode)) %>% 
  select(drv)
```

Success!!!